---
title: "Lab 3_2"
author: "Yi Ren"
date: "1/25/2019"
output: 
  html_document: 
    keep_md: yes
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
getwd()
```

#load the tidyverse library
```{r}
library("tidyverse")
```

#read the fish data
```{r}
fish <- readr::read_csv("~/Desktop/class_files-master/Gaeta_etal_CLC_data.csv")
```

#can view the file
```{r}
view(fish)
```

##summary functions

#names command: outputs the colomn names
```{r}
names(fish)
```
#head command: outputs the first 6
```{r}
head(fish)
```
#summary command: outputs summary
```{r}
summary(fish)
```
#glimpse command
```{r}
glimpse(fish)
```
#str command
```{r}
str(fish)
```


## Practice
1. Load the data `mammal_lifehistories_v2.csv` and place it into a new object called `mammals`.
```{r}
mammals <- readr::read_csv("~/Desktop/class_files-master/mammal_lifehistories_v2.csv")
```
2. Provide the dimensions of the data frame.
# 1440 observations of 13 variables

3. Display the column names in the data frame. 
```{r}
names(mammals)
```

4. Use str() to show the structure of the data frame and its individual columns; compare this to glimpse(). 
```{r}
str(mammals)
```

```{r}
glimpse(mammals)
```


5. Print out the first few rows of the data using the function head().  
```{r}
head(mammals)
```


##Dplyr

#filter():looks through a data frame, pulls out data that meet the entered criteria
```{r}
names(fish)
```

```{r}
filter(fish,lakeid=="AL")
```
#filter: can use >, >=, ==, &(and), |(or), etc.
```{r}
filter(fish,length > 350 & lakeid =="AL")
```

```{r}
filter(fish, length == 167 | length == 175)
```
# ! means not
```{r}
filter(fish, !length == 175)
```


##practice

1. Filter the `fish` data to include the samples from lake `DY`.
```{r}
filter(fish, lakeid == "DY")
```

2. Filter the data to include all lakes except AL.
```{r}
filter(fish, !lakeid == "AL")
```

3. Filter the data to include all lakes except AL and DY.
```{r}
filter(fish, !lakeid == "AL" & !lakeid == "DY")
```

4. Filter the data to include all fish with a scale length greater than or equal to 11.
```{r}
filter(fish, scalelength >= 11)
```

5. Filter the data to include fish only from lake AL and with a scalelength greater than or equal to 2 and less than or equal to 4.
```{r}
filter(fish, lakeid == "AL",
       scalelength >= 2 & scalelength <= 4)
```


## Pipes
#shortcut: shift + command + m
#pipe the data into the code to make it look cleaner
```{r}
fish %>%
  filter(lakeid == "AL")
```

## Select()
#filter out by rows
```{r}
fish %>% 
  select(lakeid, scalelength)
```

#combine filter & select
```{r}
fish %>% 
  select(lakeid, scalelength) %>% 
  filter(lakeid == "AL",
       scalelength >= 2 & scalelength <= 4)
```
