---
title: "Lab 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


#If a vector contains multiple types of data, it changes all data into character. Could cause problems in the future. So just keep vector in one data type.
```{r}
new_vector <- c (1,2,3,"orange","blue")
new_vector
class(new_vector)
```

```{r}
options(scipen=999) #disables scientific notation
organism<-c("Human","Mouse","Fruit Fly", "Roundworm","Yeast")
genomeSizeBP<-c(3000000000,3000000000,135600000,97000000,12100000)
estGeneCount<-c(30000,30000,13061,19099,6034)
```

We can combine them into a single vector.
```{r}
genome_size <- c(genomeSizeBP, estGeneCount)
genome_size
```

And provide names.
```{r}
names(genome_size) <- c("genome size BP", "genome size BP", "genome size BP", "genome size BP", "genome size BP", "gene count", "gene count", "gene count", "gene count", "gene count")
genome_size
```

You can also select elements of a vector using `[]`
```{r}
humans <- genome_size[c(1, 6)]
humans
```

**Data Matrices.** Vectors may seem inefficient, but they are a fundamental data structure in R. A data matrix is a little better at organizing the data.
```{r}
genome_matrix <- matrix(genome_size, nrow=2, byrow=T)
genome_matrix
```

Let's name the columns and rows.
```{r}
colnames(genome_matrix) <- organism
rownames(genome_matrix) <- c("genome size BP", "gene count")
genome_matrix
```

Once we have the matrix, we can use functions to make some calculations.
```{r}
means <- rowMeans(genome_matrix)
```

Add them to the matrix
```{r}
genome_matrix2 <- cbind(genome_matrix, means)
genome_matrix2
```


-------------------------------
*FOCUS OF TODAY: DATA FRAME
data frame: like a data table in excel, but every column is a vector

6. **Data Frames.**Most data that you will encounter will be structured as a data matrix. Notice that each column is a vector!

#assign a name to each of the column
```{r}
genome_frame <- data.frame(organism=organism, genomeSizeBP=genomeSizeBP, estGeneCount=estGeneCount)
genome_frame
```

